Efficient Compilation of High Level Python Numerical Programs with Pythran

### 2014-08-25 doc/papers/cgo2014/paper.tex
Introduction

\cite{rossum97, scipy}

%%
%%
\section{A Crowd-Sourced Numpy Benchmark}

@@

%%
%%
\section{Optimizations Opportunities in a Typical Numpy Kernel}

@@

%%
%%
\section{Existing Compilation Approach}

@@

%%
%%
\section{Compiler - Runtime Cooperation}

@@

%%
%%
\section{Experiments}

@@

\subsection{Expermimental Setup}

>>>>

The Python language~\cite{rossum97} now provides a full toolkit to carry out
scientific experiments~\cite{scipy}, from core scientific routines with the
Numpy package, to scientific packages with the Scipy package, plotting
facilities with the Matplotlib package, enhanced terminal and notebooks with
IPython. As a consequence, there has been a move from historical languages like
Fortran to Python, as showcased by the success of the Scipy conference.

As Python based scientific tools gets widely used, the question of High
performance Computing naturally arises, and it is the focus of many recent
research. Indeed, although there is a great gain in productivity when using
these tools, there is also a performance gap that needs to be filled.

This paper focuses on compilation techniques that are relevant for the
optimization of high-level numerical kernels written in Python using the Numpy
package. Section~\ref{sec:stackoverflow} presents a case study based on user
input gathered on the stackoverflow question and answer site.
Section~\ref{sec:optimize} focuses on a simple yet representative kernel and
showcases the optimization opportunities typically found in Numpy-based
kernels. Existing compilation approach to optimize such kernels are discussed
in Section~\ref{sec:compilers}, then Section~\ref{sec:pythran} emphasises on a
compiler-runtime cooperative approach to solve the performance issues.
Experimentations showcasing important performance gains are detailed in
Section~\ref{sec:xp}.

%%
%%
\section{A Crowd-Sourced Numpy Benchmark}
\label{sec:stackoverflow}

@@

The benchmark is publicly available on
\url{https://github.com/serge-sans-paille/numpy-benchmarks}.

%%
%%
\section{Optimizations Opportunities in a Typical Numpy Kernel}
\label{sec:optimize}

@@

%%
%%
\section{Existing Compilation Approach}
\label{sec:compilers}

@@

%%
%%
\section{Compiler - Runtime Cooperation}
\label{sec:pythran}

@@

%%
%%
\section{Experiments}
\label{sec:xp}

@@

\subsection{Experimental Setup}

\subsection{Performance of the rosenbrock Function}
